---
title: "Reproducibe Research Proyect 1"
author: "Abner Aranda"
date: "8/9/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading and preparing data
```{r}
if (!file.exists("activity.csv") )
    {
     dlurl <- 'http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
     download.file(dlurl,destfile='repdata%2Fdata%2Factivity.zip',mode='wb')  
     unzip('repdata%2Fdata%2Factivity.zip')
    }
activity <- read.csv("activity.csv")
```

What is the mean total number of steps taken per day?
We start by creating a histogram of steps per day
```{r}
Act_without_na <- subset(activity, !is.na(activity$steps))
step_per_day <- aggregate(steps ~ date, Act_without_na, sum)
hist(step_per_day $steps, breaks = 53, col = "green", xlab = "Number of steps", main = "Histogram of the total number of steps taken each day")
```
Calculate mean and median of the total number of steps taken per day
Mean
```{r}
act_mean <- mean(step_per_day$steps)
print(act_mean)
```
Median
```{r}
act_med <- median(step_per_day$steps)
print(act_med)
```

What is the average daily activity pattern?
Plot of steps per interval
```{r}
steps_per_interval <- aggregate(steps ~ interval, Act_without_na, mean)
plot(steps_per_interval$interval, steps_per_interval$steps, type="l", xlab="Interval", ylab="Number of Steps", main="Average Number of Steps per Day by Interval")
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_int <- steps_per_interval[which.max(steps_per_interval$steps),1]
print(max_int)

```
How many steps does that interval had?
```{r}
max_int_numb <- steps_per_interval[steps_per_interval$interval == max_int,2]
print(max_int_numb)
```

Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum_NA <- sum(!complete.cases(activity))
sum_NA
```
Create a new dataset that is equal to the original dataset but with the missing data filled in. Create a histogram
```{r}
na_index <- which(is.na(as.character(activity$steps)))
complete_act <- activity
complete_act[na_index, ]$steps <- unlist(lapply(na_index, FUN=function(na_index){steps_per_interval[activity[na_index, ]$interval== steps_per_interval$interval, ]$steps}))
step_per_day_complete <- aggregate(steps ~ date, data = complete_act, sum)
hist(step_per_day_complete $steps, breaks = 53, col = "green", xlab = "Number of steps", main = "Histogram of the total number of steps taken each day")
```
Mean
```{r}
mean(step_per_day_complete$steps)
```
Median
```{r}
median(step_per_day_complete$steps)
```
Both mean and median has little to no changecompared with the incomplete data

Are there differences in activity patterns between weekdays and weekends?
Lets separate the data between weekend and weekdays
```{r}
complete_act$date <- as.Date(complete_act$date, format = "%Y-%m-%d")
weekdays <- c("Monday","Tuesday","Wednesday","Thursday","Friday")
complete_act $day_week = as.factor(ifelse(is.element(weekdays(as.Date(complete_act $date)),weekdays), "Weekday", "Weekend"))
steps_by_day_type <- aggregate(steps ~ interval + day_week, complete_act, mean)
library(lattice)
xyplot(steps_by_day_type$steps ~ steps_by_day_type$interval| steps_by_day_type$day_week, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```
The weekends do have more activity, although weekdays have the biggest peak of activity, pressumible during the morning